---
title: "account_phenos"
author: "Ashley Mulford"
date: "November 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
```

#Analysis of covariates and phenotypes

##Two NOAC Drugs: Apixaban and Rivaroxiban

##### Sample Sizes:
##### Apixaban: 142 total, 125 with AFX, 141 with CrCl
##### Rivaroxaban: 110 total, 102 with AFX, 107 with CrCl

```{r}
noac_data <- fread("account_noac_processed.csv", header = T)

#subset by drug
apixaban <- subset(noac_data, `Which NOAC Drug Used`  == "apixaban")
rivaroxaban <- subset(noac_data, `Which NOAC Drug Used`  == "rivaroxaban")

#remove individuals with missing Anti-factor Xa activity data
api_no_miss_AFX<-subset(apixaban, `Anti-factor Xa activity (ng/mL)` != "MD")
riv_no_miss_AFX<-subset(rivaroxaban, `Anti-factor Xa activity (ng/mL)` != "MD")

#remove individuals with missing Creatinine clearance data
api_no_miss_CrCl<-subset(apixaban, `Creatinine clearance (CrCl)` != "MD")
riv_no_miss_CrCl<-subset(rivaroxaban, `Creatinine clearance (CrCl)` != "MD")

```

### Anti-factor Xa Activity Distributions:

#### Apixaban:

```{r}
shapiro.test(as.integer(api_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`))

hist(as.integer(api_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`), main = "Distribution of Anti-Factor Xa Activity - Apixaban", xlab = "Anti-Factor Xa Activity (ng/mL)")
```

##### Anti-Factor Xa Activity Distribution for Apixaban is not normal.


#### Rivaroxaban:

```{r}
shapiro.test(as.integer(riv_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`))

hist(as.integer(riv_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`), main = "Distribution of Anti-Factor Xa Activity - Rivaroxaban", xlab = "Anti-Factor Xa Activity (ng/mL)")
```

##### Anti-Factor Xa Activity Distribution for Rivaroxaban is not normal.


### Dosage vs Anti-factor Xa activity:

#### Apixaban:

```{r}

api_no_miss_AFX_subset<-api_no_miss_AFX[, c(1:20,81)]

api_no_miss_AFX_dose5<-subset(api_no_miss_AFX_subset, `Therapeutic dose (mg/day)` == "5")
api_no_miss_AFX_dose10<-subset(api_no_miss_AFX_subset, `Therapeutic dose (mg/day)` == "10")

t.test(as.integer(api_no_miss_AFX_dose5$`Anti-factor Xa activity (ng/mL)`), as.integer(api_no_miss_AFX_dose10$`Anti-factor Xa activity (ng/mL)`))

ggplot(api_no_miss_AFX_subset, aes(x=as.integer(api_no_miss_AFX_subset$`Therapeutic dose (mg/day)`), y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Apixaban Dose (mg/day)") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Anti-factor Xa activity - Apixaban")

summary(lm(as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`)~as.integer(api_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)))

```

##### Anti-Factor Xa Activity is significantly different (increased) for Apixaban dose of 10 than of 5. 


#### Rivaroxaban:

```{r}

riv_no_miss_AFX_subset<-riv_no_miss_AFX[, c(1:20,81)]

dose_10<-c()

for(i in 1:102) {
  if (as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)[i] == 10) {
    dose_10<-c(dose_10, 1)
  }
  else {
    dose_10<-c(dose_10, 0)
  }
}

dose_15<-c()

for(i in 1:102) {
  if (as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)[i] == 15) {
    dose_15<-c(dose_15, 1)
  }
  else {
    dose_15<-c(dose_15, 0)
  }
}

dose_20<-c()

for(i in 1:102) {
  if (as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)[i] == 20) {
    dose_20<-c(dose_20, 1)
  }
  else {
    dose_20<-c(dose_20, 0)
  }
}

dose_30<-c()

for(i in 1:102) {
  if (as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)[i] == 30) {
    dose_30<-c(dose_30, 1)
  }
  else {
    dose_30<-c(dose_30, 0)
  }
}

anova_dosages_AFX<-cbind(dose_10,dose_15,dose_20,dose_30)

#ANOVA
anova(lm(as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`) ~ anova_dosages_AFX))

ggplot(riv_no_miss_AFX_subset, aes(x=as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`), y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Rivaroxaban Dose (mg/day)") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Anti-factor Xa activity - Rivaroxaban")

summary(lm(as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`)~as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`)))

```

##### Differences in Rivaroxaban dosage have a significant effect on Anti-factor Xa activity.

### Age vs Anti-factor Xa activity:

#### Apixaban:

```{r}

ggplot(api_no_miss_AFX_subset, aes(x=as.integer(api_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Anti-factor Xa activity - Apixaban")

summary(lm(as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`)~as.integer(api_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`)))

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_AFX_subset, aes(x=as.integer(riv_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Anti-factor Xa activity - Rivaroxaban")

summary(lm(as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`)~as.integer(riv_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`)))

```

### Gender vs Anti-factor Xa activity:

#### Apixaban:

```{r}

api_no_miss_AFX_genF<-subset(api_no_miss_AFX_subset, `Gender` == "F")
api_no_miss_AFX_genM<-subset(api_no_miss_AFX_subset, `Gender` == "M")

t.test(as.integer(api_no_miss_AFX_genF$`Anti-factor Xa activity (ng/mL)`), as.integer(api_no_miss_AFX_genM$`Anti-factor Xa activity (ng/mL)`))

ggplot(api_no_miss_AFX_subset, aes(x=api_no_miss_AFX_subset$Gender, y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  geom_boxplot() +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Anti-factor Xa activity - Apixaban")

```

##### Anti-Factor Xa Activity is significantly different (increased) in women compared to men taking Apixaban


#### Rivaroxaban:

```{r}

riv_no_miss_AFX_genF<-subset(riv_no_miss_AFX_subset, `Gender` == "F")
riv_no_miss_AFX_genM<-subset(riv_no_miss_AFX_subset, `Gender` == "M")

t.test(as.integer(riv_no_miss_AFX_genF$`Anti-factor Xa activity (ng/mL)`), as.integer(riv_no_miss_AFX_genM$`Anti-factor Xa activity (ng/mL)`))

ggplot(riv_no_miss_AFX_subset, aes(x=riv_no_miss_AFX_subset$Gender, y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#3399ff") +
  geom_boxplot() +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Anti-factor Xa activity - Rivaroxaban")

```

##### Anti-Factor Xa Activity is significantly different (increased) in women compared to men taking Rivaroxaban


#### Conclusion:
##### For both Apixaban and Rivaroxaban: Anti-Factor Xa Activity is not normal distributed and will need to be rank-normalized; dosage, age, and gender will need to be used as covariates.




### Creatinine clearance Distributions:

#### Apixaban:

```{r}
shapiro.test(as.integer(api_no_miss_CrCl$`Anti-factor Xa activity (ng/mL)`))

hist(as.integer(api_no_miss_CrCl$`Creatinine clearance (CrCl)`), main = "Distribution of Creatinine clearance - Apixaban", xlab = "Creatinine clearance (CrCl)")
```

##### Creatinine Clearance Distribution for Apixaban is not normal.


#### Rivaroxaban:

```{r}
shapiro.test(as.integer(riv_no_miss_CrCl$`Anti-factor Xa activity (ng/mL)`))

hist(as.integer(riv_no_miss_CrCl$`Creatinine clearance (CrCl)`), main = "Distribution of Creatinine clearance - Rivaroxaban", xlab = "Creatinine clearance (CrCl)")
```

##### Creatinine Clearance Distribution for Rivaroxaban is not normal.


### Dosage vs Creatinine clearance:

#### Apixaban:

```{r}

api_no_miss_CrCl_subset<-api_no_miss_CrCl[, c(1:20,77)]

api_no_miss_CrCl_dose5<-subset(api_no_miss_CrCl_subset, `Therapeutic dose (mg/day)` == "5")
api_no_miss_CrCl_dose10<-subset(api_no_miss_CrCl_subset, `Therapeutic dose (mg/day)` == "10")

t.test(as.integer(api_no_miss_AFX_dose5$`Anti-factor Xa activity (ng/mL)`), as.integer(api_no_miss_AFX_dose10$`Anti-factor Xa activity (ng/mL)`))

ggplot(api_no_miss_CrCl_subset, aes(x=as.integer(api_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`), y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Apixaban Dose (mg/day)") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Creatinine clearance - Apixaban")

summary(lm(as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`)~as.integer(api_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`)))  

```

##### Creatinine Clearance is significantly different (increased) for Apixaban dose of 10 than of 5. 

#### Rivaroxaban:

```{r}

riv_no_miss_CrCl_subset<-riv_no_miss_CrCl[, c(1:20,77)]

#ANOVA

dose_10<-c()

for(i in 1:107) {
  if (as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`)[i] == 10) {
    dose_10<-c(dose_10, 1)
  }
  else {
    dose_10<-c(dose_10, 0)
  }
}

dose_15<-c()

for(i in 1:107) {
  if (as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`)[i] == 15) {
    dose_15<-c(dose_15, 1)
  }
  else {
    dose_15<-c(dose_15, 0)
  }
}

dose_20<-c()

for(i in 1:107) {
  if (as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`)[i] == 20) {
    dose_20<-c(dose_20, 1)
  }
  else {
    dose_20<-c(dose_20, 0)
  }
}

dose_30<-c()

for(i in 1:107) {
  if (as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`)[i] == 30) {
    dose_30<-c(dose_30, 1)
  }
  else {
    dose_30<-c(dose_30, 0)
  }
}

anova_dosages_CrCl<-cbind(dose_10,dose_15,dose_20,dose_30)

#ANOVA
anova(lm(as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`) ~ anova_dosages_CrCl))


ggplot(riv_no_miss_CrCl_subset, aes(x=as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`), y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Rivaroxaban Dose (mg/day)") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Creatinine clearance - Rivaroxaban")

summary(lm(as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`)~as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`))) 

```

##### Differences in Rivaroxaban dosage have a significant effect on Creatinine Clearance. 


### Age vs Creatinine clearance:

#### Apixaban:

```{r}

ggplot(api_no_miss_CrCl_subset, aes(x=as.integer(api_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Creatinine clearance - Apixaban")

summary(lm(as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`)~as.integer(api_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`))) 

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_CrCl_subset, aes(x=as.integer(riv_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Creatinine clearance - Rivaroxaban")

summary(lm(as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`)~as.integer(riv_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`))) 

```

### Gender vs Creatinine clearance:

#### Apixaban:

```{r}

api_no_miss_CrCl_genF<-subset(api_no_miss_CrCl_subset, `Gender` == "F")
api_no_miss_CrCl_genM<-subset(api_no_miss_CrCl_subset, `Gender` == "M")

t.test(as.integer(api_no_miss_CrCl_genF$`Creatinine clearance (CrCl)`), as.integer(api_no_miss_CrCl_genM$`Creatinine clearance (CrCl)`))

ggplot(api_no_miss_CrCl_subset, aes(x=api_no_miss_CrCl_subset$Gender, y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  geom_boxplot() +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Creatinine clearance - Apixaban")

```

##### Creatinine Clearance is not significantly different in women compared to men taking Apixaban


#### Rivaroxaban:

```{r}

riv_no_miss_CrCl_genF<-subset(riv_no_miss_CrCl_subset, `Gender` == "F")
riv_no_miss_CrCl_genM<-subset(riv_no_miss_CrCl_subset, `Gender` == "M")

t.test(as.integer(riv_no_miss_CrCl_genF$`Creatinine clearance (CrCl)`), as.integer(riv_no_miss_CrCl_genM$`Creatinine clearance (CrCl)`))

ggplot(riv_no_miss_CrCl_subset, aes(x=riv_no_miss_CrCl_subset$Gender, y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#3399ff") +
  geom_boxplot() +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Creatinine clearance - Rivaroxaban")

```

##### Creatinine Clearance is significantly different (increased) in women compared to men taking Rivaroxaban

#### Conclusion:
##### For both Apixaban and Rivaroxaban: Creatinine Clearance is not normal distributed and will need to be rank-normalized; dosage, age, and gender will need to be used as covariates.




## Clopidogrel:
##### Sample Size: 163 total, 147 with Plat

```{r}

#clopidogrel dosage is 75 mg/day for all individuals

clop_data <- fread("account_clopidogrel_processed.csv", header = T)

#remove individuals with missing platelet count data
clop_no_miss_plat<-subset(clop_data, `VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU` != "MD")

clop_no_miss_plat_subset<-clop_no_miss_plat[-1, c(1:26,201)]

```

### Platelet Count Distribution:

```{r}

shapiro.test(as.integer(clop_no_miss_plat_subset$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`))

hist(as.integer(clop_no_miss_plat_subset$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`), main = "Distribution of Platelet count - Clopidogrel", xlab = "VerifyNow ADP stimulated Aggregation PRU")
```

##### Platelet Count Distribution for Clopidogrel is not normal.

### Age vs Platelet Count:

```{r}

ggplot(clop_no_miss_plat_subset, aes(x=as.integer(clop_no_miss_plat_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(clop_no_miss_plat_subset$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`))) + 
  geom_point(color = "#3399ff") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "VerifyNow ADP stimulated Aggregation PRU") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Platelet Count - Clopidogrel")

summary(lm(as.integer(clop_no_miss_plat_subset$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`)~as.integer(clop_no_miss_plat_subset$`Age at enrollment (with over 89 binning)`))) 

```

### Gender vs Platelet Count:

```{r}

clop_no_miss_plat_genF<-subset(clop_no_miss_plat_subset, `Gender` == "F")
clop_no_miss_plat_genM<-subset(clop_no_miss_plat_subset, `Gender` == "M")

t.test(as.integer(clop_no_miss_plat_genF$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`), as.integer(clop_no_miss_plat_genM$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`))

ggplot(clop_no_miss_plat_subset, aes(x=clop_no_miss_plat_subset$Gender, y=as.integer(clop_no_miss_plat_subset$`VerifyNow ADP stimulated Aggregation (while on maintenance dose of Clopidogrel) PRU`))) + 
  geom_point(color = "#3399ff") +
  geom_boxplot() +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#ff6688") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "VerifyNow ADP stimulated Aggregation PRU") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Platelet Count - Clopidogrel")

```

##### Platelet Count is not significantly different in women compared to men taking Clopidogrel


#### Conclusion:
##### For Clopidogrel: Platelet Count is not normal distributed and will need to be rank-normalized; age and gender will need to be used as covariates; dosage is constant in all patients.




