---
title: "account_phenos"
author: "Ashley Mulford"
date: "November 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
```

#Analysis of covariates and phenotypes

##Two NOAC Drugs: Apixaban and Rivaroxiban

```{r}
noac_data <- fread("account_noac_processed.csv", header = T)

#subset by drug
apixaban <- subset(noac_data, `Which NOAC Drug Used`  == "apixaban")
rivaroxaban <- subset(noac_data, `Which NOAC Drug Used`  == "rivaroxaban")

#remove individuals with missing Anti-factor Xa activity data
api_no_miss_AFX<-subset(apixaban, `Anti-factor Xa activity (ng/mL)` != "MD")
riv_no_miss_AFX<-subset(rivaroxaban, `Anti-factor Xa activity (ng/mL)` != "MD")

#remove individuals with missing Creatinine clearance data
api_no_miss_CrCl<-subset(apixaban, `Creatinine clearance (CrCl)` != "MD")
riv_no_miss_CrCl<-subset(rivaroxaban, `Creatinine clearance (CrCl)` != "MD")

```

### Anti-factor Xa Activity Distributions:

#### Apixaban:

```{r}
hist(as.integer(api_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`), main = "Distribution of Anti-Factor Xa Activity - Apixaban", xlab = "Anti-Factor Xa Activity (ng/mL)")
```

#### Rivaroxaban:

```{r}
hist(as.integer(riv_no_miss_AFX$`Anti-factor Xa activity (ng/mL)`), main = "Distribution of Anti-Factor Xa Activity - Rivaroxaban", xlab = "Anti-Factor Xa Activity (ng/mL)")
```

### Dosage vs Anti-factor Xa activity:

#### Apixaban:

```{r}

api_no_miss_AFX_subset<-api_no_miss_AFX[, c(1:20,81)]

ggplot(api_no_miss_AFX_subset, aes(x=as.integer(api_no_miss_AFX_subset$`Therapeutic dose (mg/day)`), y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  scale_x_continuous(name = "Apixaban Dose (mg/day)") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Anti-factor Xa activity - Apixaban")
  

```

#### Rivaroxaban:

```{r}

riv_no_miss_AFX_subset<-riv_no_miss_AFX[, c(1:20,81)]

ggplot(riv_no_miss_AFX_subset, aes(x=as.integer(riv_no_miss_AFX_subset$`Therapeutic dose (mg/day)`), y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Rivaroxaban Dose (mg/day)") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Anti-factor Xa activity - Rivaroxaban")



```

### Age vs Anti-factor Xa activity:

#### Apixaban:

```{r}

ggplot(api_no_miss_AFX_subset, aes(x=as.integer(api_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Anti-factor Xa activity - Apixaban")

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_AFX_subset, aes(x=as.integer(riv_no_miss_AFX_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Anti-factor Xa activity - Rivaroxaban")

```

### Gender vs Anti-factor Xa activity:

#### Apixaban:

```{r}

ggplot(api_no_miss_AFX_subset, aes(x=api_no_miss_AFX_subset$Gender, y=as.integer(api_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Anti-factor Xa activity - Apixaban")

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_AFX_subset, aes(x=riv_no_miss_AFX_subset$Gender, y=as.integer(riv_no_miss_AFX_subset$`Anti-factor Xa activity (ng/mL)`))) + 
  geom_point(color = "#ec328c") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Anti-factor Xa activity (ng/mL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Anti-factor Xa activity - Rivaroxaban")

```



###  Creatinine clearance Distributions:

#### Apixaban:

```{r}
hist(as.integer(api_no_miss_CrCl$`Creatinine clearance (CrCl)`), main = "Distribution of Creatinine clearance - Apixaban", xlab = "Creatinine clearance (CrCl)")
```

#### Rivaroxaban:

```{r}
hist(as.integer(riv_no_miss_CrCl$`Creatinine clearance (CrCl)`), main = "Distribution of Creatinine clearance - Rivaroxaban", xlab = "Creatinine clearance (CrCl)")
```

### Dosage vs Creatinine clearance:

#### Apixaban:

```{r}

api_no_miss_CrCl_subset<-api_no_miss_CrCl[, c(1:20,77)]

ggplot(api_no_miss_CrCl_subset, aes(x=as.integer(api_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`), y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Apixaban Dose (mg/day)") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Creatinine clearance - Apixaban")
  

```

#### Rivaroxaban:

```{r}

riv_no_miss_CrCl_subset<-riv_no_miss_CrCl[, c(1:20,77)]

ggplot(riv_no_miss_CrCl_subset, aes(x=as.integer(riv_no_miss_CrCl_subset$`Therapeutic dose (mg/day)`), y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Rivaroxaban Dose (mg/day)") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Dose vs Creatinine clearance - Rivaroxaban")



```

### Age vs Creatinine clearance:

#### Apixaban:

```{r}

ggplot(api_no_miss_CrCl_subset, aes(x=as.integer(api_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Creatinine clearance - Apixaban")

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_CrCl_subset, aes(x=as.integer(riv_no_miss_CrCl_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Creatinine clearance - Rivaroxaban")

```

### Gender vs Creatinine clearance:

#### Apixaban:

```{r}

ggplot(api_no_miss_CrCl_subset, aes(x=api_no_miss_CrCl_subset$Gender, y=as.integer(api_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Creatinine clearance - Apixaban")

```

#### Rivaroxaban:

```{r}

ggplot(riv_no_miss_CrCl_subset, aes(x=riv_no_miss_CrCl_subset$Gender, y=as.integer(riv_no_miss_CrCl_subset$`Creatinine clearance (CrCl)`))) + 
  geom_point(color = "#ec328c") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Creatinine clearance (CrCl)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Creatinine clearance - Rivaroxaban")

```



## Clopidogrel:

```{r}

#clopidogrel dosage is 75 mg/day for all individuals

clop_data <- fread("account_clopidogrel_processed.csv", header = T)

#remove individuals with missing platelet count data
clop_no_miss_plat<-subset(clop_data, `Platelet count (cells/microL)` != "MD")

clop_no_miss_plat_subset<-clop_no_miss_plat[-1, c(1:26,208)]

```

### Platelet Count Distribution:

```{r}
hist(as.integer(clop_no_miss_plat_subset$`Platelet count (cells/microL)`), main = "Distribution of Platelet count - Clopidogrel", xlab = "Platelet count (cells/microL) (*10^3 AuL)")
```

### Age vs Platelet Count:

```{r}

ggplot(clop_no_miss_plat_subset, aes(x=as.integer(clop_no_miss_plat_subset$`Age at enrollment (with over 89 binning)`), y=as.integer(clop_no_miss_plat_subset$`Platelet count (cells/microL)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_continuous(name = "Age") + 
  scale_y_continuous(name = "Platelet count (cells/microL) (*10^3 AuL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Age vs Platelet Count - Clopidogrel")

```

### Gender vs Platelet Count:

```{r}

ggplot(clop_no_miss_plat_subset, aes(x=clop_no_miss_plat_subset$Gender, y=as.integer(clop_no_miss_plat_subset$`Platelet count (cells/microL)`))) + 
  geom_point(color = "#ec328c") +
  stat_smooth(method="lm", se = TRUE, fullrange = TRUE, color = "#8c1788") +
  scale_x_discrete(name = "Gender") + 
  scale_y_continuous(name = "Platelet count (cells/microL) (*10^3 AuL)") +
  theme_bw() + 
  theme(text = element_text(size = 12), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gender vs Platelet Count - Clopidogrel")

```



